---
title: "Drivers of sample number"
author: "Engagement team - Freshawater Hackathon"
date: "25-27 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Introduction

We want to know what the drivers of more samples is. We think this might be driven by things like the type of training they got, the amount and types of social engagement, length of involvement, etc.

# Data

The data were provided in .csv format. These were read in and some columns were fixed. The data was saved then as a `.rdata` which we read in here

```{r read_data}
rm(list = ls())
load('data/new_data2.rdata')
library(lme4)
library(ggplot2)
library(reshape2)
library(pscl)
library(glmm)
library(sjPlot)
library(sjmisc)

# merge in the social metrics
social_metrics <- read.csv('data/Skill&com_score.csv')

new_data <- merge(x = new_data, y = social_metrics, by = 'uid', all = TRUE)

# Merge in the country data
countries <- read.csv(file = 'data/Country.csv')

new_data <- merge(x = new_data, y = countries, by = 'uid', all = TRUE)
```

Now lets do some correlative analyses

```{r first_stats}
attach(new_data)

hist(new_data$Sample, breaks = 100)

# We have a zero inflation problem

# How much data do we have for the covariates of interest?
nrow(new_data)
summary(new_data[, c('ActivePeriod', 'Paid','Staff','PCA1_Difficulty', 'WQS', 'Com_score', 'Skill_score')])
# We are going to lose lots of data through NAs in the Paid column
# WQS would be great but there are so many NAs we just can't use it

# We might be best loging some of our predictor variables
hist(log(new_data$ActivePeriod))
hist(new_data$PCA1_Difficulty)
hist(log(new_data$Com_score))
hist(log(new_data$Skill_score))

# m1 <- zeroinfl(Sample ~ log(ActivePeriod) + Paid + Staff + PCA1_Difficulty + log(Com_score) + log(Skill_score), data = new_data)
# 
# summary(m1)

# We think that the 0s in active days and in samples is confusing this analysis
# I also remove staff
# new_data_noZeros <- new_data[new_data$Sample > 0 & new_data$ActivePeriod > 0 & new_data$Staff == 0, ]

m2 <- glm(Sample ~ log(ActivePeriod+1) + Paid + PCA1_Difficulty + log(Com_score+1) + log(Skill_score+1), data = new_data, family = 'poisson')

summary(m2)

# We need to move to a mixed effects model because we want to put country in there
m3 <- glmer(Sample ~ log(ActivePeriod+1) + Paid + PCA1_Difficulty + log(Com_score+1) + log(Skill_score+1) +(1|Country), data = new_data, family = 'poisson')

summary(m3)
# Adding country as a random effect has improved the model
sjp.glmer(m3)

sjp.glmer(m3, type = "eff",
          axis.lim = list(c(0,50), c(0,5), c(0,5), c(0,5)),
          facet.grid = FALSE,
          show.p = TRUE,
          show.ci = TRUE)
```

# Results

There is a diffrence in paid and unpaid traing and how long they stay

```{r effect_of_pay}
pay <- new_data[!is.na(new_data$Paid),]
p <- ggplot(pay, aes(as.factor(Paid), na.omit(log(ActivePeriod)+1)))
p + geom_boxplot()

t1a <- t.test(x = log(na.omit(new_data$ActivePeriod[new_data$Paid == 1]+1)),
             y = log(na.omit(new_data$ActivePeriod[new_data$Paid == 0]+1)))

cat('mean of people paid to train', as.numeric(exp(t1a$estimate[1])), '\n')
cat('mean of people not paid to train', as.numeric(exp(t1a$estimate[2])), '\n')

# Lets to this again without staff or people who have provided 0 samples
pay2 <- new_data[!is.na(new_data$Paid) & new_data$Sample > 0 & new_data$Staff == 0,]
p <- ggplot(pay2, aes(as.factor(Paid), na.omit(log(ActivePeriod)+1)))
p + geom_boxplot()

t1b <- t.test(x = log(na.omit(pay2$ActivePeriod[pay2$Paid == 1]+1)),
             y = log(na.omit(pay2$ActivePeriod[pay2$Paid == 0]+1)))

cat('mean of people paid to train', as.numeric(exp(t1b$estimate[1])), '\n')
cat('mean of people not paid to train', as.numeric(exp(t1b$estimate[2])), '\n')

# So this means the true effect is something to with 0 samplers
# Who are these people?

zeros <- data.frame(p = c('no records', 'some records'),
                   paid = c(nrow(new_data[new_data$Paid == 1 & new_data$Sample == 0,]),
                            nrow(new_data[new_data$Paid == 1 & new_data$Sample > 0,])),
                   unpaid = c(nrow(new_data[new_data$Paid == 0 & new_data$Sample == 0,]),
                              nrow(new_data[new_data$Paid == 0 & new_data$Sample > 0,])))
# counts to proportions
pzeros <- zeros
pzeros$paid <- zeros$paid/sum(zeros$paid)
pzeros$unpaid <- zeros$unpaid/sum(zeros$unpaid)

mzeros <- melt(pzeros)
colnames(mzeros)[3] <- 'Proportion_of_users'


ggplot(aes(x = variable,
           y = Proportion_of_users,
           fill = p),
       data = mzeros) + geom_bar(stat = 'identity', position=position_dodge())

# Looks like paid people sign up and do nothing
chisq.test(x = t(zeros[,2:3]))

# look in more detail
hist(new_data$Sample[new_data$Paid == 1], xlim = c(0,50), breaks = c(0:80, max(new_data$Sample)))
hist(new_data$Sample[new_data$Paid == 0], xlim = c(0,50), breaks = c(0:80, max(new_data$Sample)))

# Now lets do that same as this but look at people who have had training in uploading


uploaded <- data.frame(p = c('no records', 'some records'),
                   uploaded = c(nrow(new_data[new_data$Upload == 1 & new_data$Sample == 0,]),
                            nrow(new_data[new_data$Upload == 1 & new_data$Sample > 0,])),
                   not_uploaded = c(nrow(new_data[new_data$Upload == 0 & new_data$Sample == 0,]),
                              nrow(new_data[new_data$Upload == 0 & new_data$Sample > 0,])))
# counts to proportions
puploaded <- uploaded
puploaded$uploaded <- puploaded$uploaded/sum(puploaded$uploaded)
puploaded$not_uploaded <- puploaded$not_uploaded/sum(puploaded$not_uploaded)

muploaded <- melt(puploaded)
colnames(muploaded)[3] <- 'Proportion_of_users'


ggplot(aes(x = variable,
           y = Proportion_of_users,
           fill = p),
       data = muploaded) + geom_bar(stat = 'identity', position=position_dodge())

# Looks like upload training makes no difference
chisq.test(x = t(uploaded[,2:3]))
```


```{r effect_of_team}
team <- new_data[!is.na(new_data$Team) & new_data$Sample > 0,]
p <- ggplot(team, aes(as.factor(Team), na.omit(log(Sample)+1)))
p + geom_boxplot()

t2 <- t.test(x = log(na.omit(new_data$Sample[new_data$Team == 1]+1)),
             y = log(na.omit(new_data$Sample[new_data$Team == 0]+1)))

cat('mean of people in teams', as.numeric(exp(t2$estimate[1])), '\n')
cat('mean of people not in teams', as.numeric(exp(t2$estimate[2])), '\n')
```
